### 2022-02-21 research/process/acknowledgement.tex

% Copyright (c) 2021 Tobias Briones. All rights reserved.
% SPDX-License-Identifier: CC-BY-4.0
%
% This file is part of https://github.com/tobiasbriones/
% cp-unah-mm700-agricultural-soil-sampling-for-data-analysis
%
% This source code is licensed under the Creative Commons Attribution 4.0
% International License found in the LICENSE-CC file in the root directory of
% this source tree or at https://spdx.org/licenses/CC-BY-4.0

\section{Reconocimientos}

Se agradece al equipo de \textit{PL Intelligent Solutions Co} por las visiones dadas sobre el contexto e importancia del problema de investigación en el país.